---
title: "Just One Graph, How Hard Can It Be"
author: "Ben Jaletzke"
date: "2020-10-20"
summary: "Yes, this 'project' is just a simple graph. The reason I am sharing it is simple: It was the first real encounter with R I've had, and it immediately showed me that I love and hate this language."
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
```

```{r}
figure_3 <- read_csv(here::here("content/data_analytics/baby_steps/", "CDC_Males.csv"))

adj_na <- figure_3 %>% 
  filter(!is.na(gun.house.prev.category))
```

Try to guess which line caused my group and me the biggest headaches. Did you spot it? No? We spent what felt like years trying to figure out how to add the line "Spearman's correlation: 0.72" to the graph. Finally, late at night and not too long before the deadline, we decided to just add it as text. "Fine, we have been outsmarted, we cannot figure it out. There's surely a fancy method to get ggplot to show analysis output that we simply do not know yet."

Imagine my surprise when, a few days after, I found the git repo of the researcher who orignially created this graph, and she, too, used geom_text(). I felt both incredible relief and frustration. Ultimately, isn't that what programming is all about?

```{r}
library(ggrepel)
figure.3 <- ggplot(adj_na, aes(adjusted.suicide.White, adjusted.homicide.White)) + 
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), col = "black", pch = 21, alpha = 0.9) + #add the points sized by population
  geom_text_repel(aes(label = ST), size = 4, min.segment.length = 1) + #add the labels _repel limits overlap
  scale_fill_manual(values = c("#fef0d9", "#fdcc8a", "#fc8d59", "#d7301f")) + #orange colour ramp
  ylab("White homicide rate (per 100,000 per year)") +
  xlab("White suicide rate (per 100,000 per year)") + 
  scale_y_continuous(limits = c(0, 5)) +
  scale_x_continuous(limits = c(0, 30)) +
  theme_bw() +
  theme() + 
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  labels = c("200k", "500k", "1.5m", "3m", "7m"),
                  max_size = 15) +
  guides(fill = guide_legend(title = "Gun ownership", 
                             override.aes = list(size = 5), order = 1), 
         size = guide_legend(title = "White population"), order = 2) + 
  geom_text(aes(x = 25, y = 0.2, 
                label = paste0("Spearman's correlation: 0.72" #round(spearman.white$spearmans, 2)
                               )), 
            check_overlap = T)
# Output
figure.3
```




